Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd
updated seed

## Overview

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 `sample size` (sample_size: The total number of respondents participating in the poll) \
 `days_taken_from_election` (the time gap between the poll start date and the real election date i.e days_taken_from_election = real US election date - poll start date) \
 `pollscore` (A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
`state` (The U.S. state where the poll was conducted or focused) \
`methodology` (The method used to conduct the poll)\

In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships. The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity. The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering. No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.

@@

@@

@@

## Basic Statistics Summary for Data

@@

## Relationship Between Variables

@@

## Basic Model
```{r}
#| label: fig-model-pct-pollscore
#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-model-pct-pollscore-etc
#| echo: false
#| warning: false
just_harris_data = Harris_data |> na.omit()
model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
# summary(model_MLR)
just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))

predictions = predict(model_MLR, just_harris_data)
```

>>>>

@@

The raw data from FiveThirtyEight contains 52 columns, all of the column headers are displayed below: 
```{r}
#| echo: false
#| warning: false
raw_data <- read_csv(here("data/01-raw_data/president_polls.csv"))

# Get the column names of your data
column_names <- colnames(raw_data)

# Reshape the column names into a matrix with, for example, 4 columns
num_cols <- 3
column_matrix <- matrix(column_names, ncol = num_cols, byrow = TRUE)

# Convert the matrix to a data frame for kable
column_df <- as.data.frame(column_matrix)

# Display in multiple columns with kable
kable(column_df, format = "latex", booktabs = TRUE, col.names = NULL)
```

These columns can be categorized into three types, response variable, numeric predictors and categorical predictors.

The response variable is:

- pct: The percentage of support or vote share that each candidate (Kamala Harris or Donald Trump) received in the poll.

Example of numeric predictors are:

- sample_size: The total number of respondents in each poll.
- pollscore: A numeric score representing the error and bias of the pollster. Negative numbers are better.
- numeric_grade: A numeric grade assigned to each pollster, reflecting pollster quality or reliability.

Example of categorical predictors are:

- state: The U.S. state in which the poll was conducted or targeted.
- methodology: The method used to conduct the poll (e.g., online, phone, in-person).
- party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
- candidate_name: The name of the candidate, either Kamala Harris or Donald Trump.
- pollster: The name of the polling organization that conducted the poll.
- stage: The stage of the election (e.g., "general").

<!-- ## Data Exploration

In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships. The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity. The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering. No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.  -->

@@

## Cleaned data

@@

### Basic Statistics Summary for Cleaned Data

@@

### Relationship Between Variables

@@

## Basic SLR Model
```{r}
#| label: fig-model-pct-pollscore
#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-model-pct-sample_size
#| fig-cap: "Actual vs. Predicted Values for pct Using sample_size as Predictor"
#| echo: false
#| warning: false
lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
predictions = predict(lm_model1, just_harris_data)
# summary(lm_model1)

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()

In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor. This indicate that sample_size has limited impact on pct.

## MLR model
```{r}
#| label: fig-model-pct-pollscore-etc
#| echo: false
#| warning: false
just_harris_data = Harris_data |> na.omit()
model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
# summary(model_MLR)
just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))

predictions = predict(model_MLR, just_harris_data)
